---
title: "OmicsAtlas Bulk RNA-seq Report"
format:
  html:
    toc: true
    embed-resources: true
params:
  report_data_path: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
report_data <- readRDS(params$report_data_path)
deseq2_res <- report_data$deseq2_res
edger_res <- report_data$edger_res
comparison <- report_data$comparison
enrichment_deseq2 <- report_data$enrichment_deseq2
enrichment_edger <- report_data$enrichment_edger
deconvolution_props <- report_data$deconvolution_props
dataset_label <- report_data$dataset_label
```

**Dataset:** `r dataset_label`

Every table, count, and plot below is computed directly from the DESeq2/edgeR/
enrichment/deconvolution objects the pipeline itself produced for this dataset
run - this report has no hardcoded numbers or narrative.

## Differential expression summary

```{r}
de_summary <- data.frame(
  method = c("DESeq2", "edgeR"),
  n_tested = c(nrow(deseq2_res), nrow(edger_res)),
  n_significant = c(sum(deseq2_res$is_significant), sum(edger_res$is_significant))
)
knitr::kable(de_summary)
```

### Top DESeq2 genes

```{r}
top_deseq2 <- deseq2_res[order(deseq2_res$padj), ]
knitr::kable(head(top_deseq2, 10))
```

### Top edgeR genes

```{r}
top_edger <- edger_res[order(edger_res$FDR), ]
knitr::kable(head(top_edger, 10))
```

## DESeq2 vs. edgeR comparison

```{r}
overlap_summary <- data.frame(
  category = c("Both methods", "DESeq2 only", "edgeR only"),
  n_genes = c(
    length(comparison$overlap$both),
    length(comparison$overlap$deseq2_only),
    length(comparison$overlap$edger_only)
  )
)
knitr::kable(overlap_summary)
```

```{r}
comparison$venn_plot
```

```{r, results = "asis"}
if (nrow(comparison$discrepancy) > 0) {
  print(knitr::kable(head(comparison$discrepancy, 20)))
} else {
  cat("No genes were called significant by exactly one method on this dataset.")
}
```

## Functional enrichment

```{r, results = "asis"}
render_enrichment_section <- function(enrichment, method_label) {
  cat(sprintf("### %s\n\n", method_label))
  if (is.null(enrichment)) {
    cat("Not run for this dataset.\n\n")
    return(invisible())
  }
  if (!is.null(enrichment$go) && nrow(as.data.frame(enrichment$go)) > 0) {
    go_df <- as.data.frame(enrichment$go)
    cat("**GO enrichment (top terms):**\n\n")
    print(knitr::kable(head(go_df[, c("Description", "GeneRatio", "p.adjust")], 10)))
    cat("\n\n")
  } else {
    cat("No significant GO enrichment.\n\n")
  }
  if (!is.null(enrichment$kegg) && nrow(as.data.frame(enrichment$kegg)) > 0) {
    kegg_df <- as.data.frame(enrichment$kegg)
    cat("**KEGG enrichment (top terms):**\n\n")
    print(knitr::kable(head(kegg_df[, c("Description", "GeneRatio", "p.adjust")], 10)))
    cat("\n\n")
  } else {
    cat("No significant KEGG enrichment (or not run).\n\n")
  }
}
render_enrichment_section(enrichment_deseq2, "DESeq2 significant genes")
render_enrichment_section(enrichment_edger, "edgeR significant genes")
```

## BayesPrism deconvolution

```{r, results = "asis"}
if (!is.null(deconvolution_props)) {
  print(knitr::kable(round(deconvolution_props, 3)))
} else {
  cat("Deconvolution was not run for this dataset.")
}
```
